---
title: "R Practice"
author: "Siddarth Ravikanti"
date: "6/2/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## Intro to markdown

Markdown allows you to do things like **bold**, *italic*, and more!

Make a list with asterisks like:

\* bold with double-asterisks

\* italics with underscores

\* code-type fond with backticks

You can also use hyphens instead of asterisks.

For a numbered list:

1\. bold...

1\. italic...

1\. code....

Hyperlinks are like [Hyperlinked text](https://mylink.org)


Load gapminder data in along with plotting package.

```{r load_data}
library(gapminder)
library(ggplot2)

gap <- gapminder


```

Try and fail to plot the data.

```{r make-plot}
plot(lifeExp ~ year, data = gap)


```

```{r}
temp <- c(1,2,3)

```


```{r make_cats}

cats <- data.frame(coat = c("calico", "black", "tabby"), weight = c(2.1, 5.0, 3.2), likes_string = c(1,0,1))

write.csv(x = cats, file = "data/feline-data.csv", row.names = FALSE)

```


```{r read_cats}
cats2 <- read.csv('data/feline-data.csv', stringsAsFactors = T)
str(cats2)


```


Creating vectors and coercing types 

```{r}
my_vector <- vector(mode = 'character', length = 3)

combine_vector <- c(1,2, "bananas")

```

change names

```{r}
names(cats)[3] <- "likes_hunting"
```


add columns and rows to a df
```{r}
cats <- read.csv('data/feline-data.csv')
age <- c(2,3,5)
cats <- cbind(cats,age)
newRow <- list("tortoiseshell",3.3,TRUE, 9)
cats <-  rbind(cats, newRow)
```

Create a vector called human_age by multiplying cats age by 7.
Convert humanage to a factor. 
convert human age back to a numeric vector by using the as.numeric() function. Now divide it by 7 to get the original ages back. 

```{r}
human_age <- cats$age * 7
human_age <- as.factor(human_age)
human_age <- as.numeric(as.character(human_age))
human_age <- human_age/7
human_age
```

remove rows and columns

```{r}
cats$coat <- as.factor(cats$coat)
cats <- rbind(cats,c("white",3.4,0,1))
#remove row

#cats <- cats[-5,]
cats <- na.omit(cats)

cats[,-4]

drop <- names(cats) %in% c("weight")
```



```{r}
library(readxl)
gapminder <- read_excel("data/gapminder_data.xlsx")
gapminder
```


Inspecting data frame
```{r}
summary(gapminder)
summary(gapminder$lifeExp)

gapminder$country <- factor(gapminder$country)
gapminder$continent <-  factor(gapminder$continent)
summary(gapminder)

head(gapminder)
tail(gapminder)
gapminder[sample(nrow(gapminder), 10),]

gapminder[1:10,2:5]

```

##Subsetting Data

```{r}
x <- c(5.4,6.2,7.1,4.8,7.5)
names(x) <-  c('a','b','c','d','e')

x[c(1,3)]
x[1:3]
x[c(1,1,3)]
x[6]
x[-2]
x[c(-1,-5)]

x <- x[-4]

x[x>7]

x[x>7 & x<7.5]
x[x>7 | x<7.5]

x[! names(x) %in% c("a", "c")]
```


Select Southeast Asian countries from gapminder

```{r}
seAsia <- c("Myanmar", "Thailand", "Cambodia", "Vietnam", "Laos")

seAsiaGap <- gapminder[gapminder$country %in% seAsia, ]

summary(seAsiaGap)
```

Extract observations from 1957

```{r}
gap57 <- gapminder[gapminder$year > 1959 &  gapminder$year > 1970,]

gap57
```


**PICTURES!!!!***

```{r}
library(ggplot2)

ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + geom_point()
```


```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) + geom_point()
```

Lines!!!

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, col = continent)) + geom_line()

```


Lines that make a little sense 
```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, by = country)) + geom_line(mapping = aes(color = continent)) + geom_point(color = "yellow")
```



```{r}
logged_points <- ggplot(gapminder, mapping = aes(x=gdpPercap, y = lifeExp)) + geom_point() + scale_x_log10() 
```

```{r}
logged_points + geom_smooth(method = "loess")
```


```{r}
americas <- gapminder[gapminder$continent == "Americas",]

ggplot(data = americas, mapping = aes(x = year, y = lifeExp)) + geom_line() + facet_wrap(~country) + theme(axis.text.x = element_text(angle = 45))

ggsave(filename = "results/lifeExp_americas.png")
```


Save Americas !!

```{r}
write.csv(file = "clean_data/americas_gap.csv", x = americas, row.names = F)
```

dplyr!!!
```{r}
mean(gapminder[gapminder$continent == "Africa", "gdpPercap"])
```

```{r}
library("dplyr")
year_country_gdp <- select(gapminder, year, country, gdpPercap)
str(year_country_gdp)
smaller_gap <- gapminder %>% select(-continent) %>% rename(gdp_per_capita = gdpPercap)
```

```{r}
y_cntr_gdp_eur <- gapminder %>% filter(continent == "Europe") %>% select(year, country, gdpPercap)
y_cntr_gdp_eur
```
combine with grouping 

```{r}
gapminder %>% group_by(continent) %>% summarize(mean_gdpPercap = mean(gdpPercap), sd_gdpPercap = sd(gdpPercap), mean_raw_gdp = mean(gdpPercap * pop), tot_raw_gdp = sum (gdpPercap*pop))
```

Mutate!

```{r}
gapminder %>% mutate(gdp_billion = gdpPercap*pop/10^9) %>% group_by(continent, year) %>% summarize(mean_gdp_bill = mean(gdp_billion))
```


combine with ggplot

```{r}
gapminder %>% filter(continent == "Americas") %>% ggplot(mapping = aes(x = year, y = lifeExp)) + geom_line() + facet_wrap(~ country)
```


